---
title: "davis-allison-ADA-homework-2"
author: "Allison Davis"
date: "February 20, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## MODULE 11
#### Probabilities and Distributions

### CHALLENGE 1:
*X = no. calls heard in each session;* *expected mean = lambda = 15*

#### 1a. Pr(X>8) in one session

``` {r}
1- ppois(q=8, lambda = 15)

```

#### 1b. Pr(X=0) in one session

```{r}
ppois(q=0, lambda = 15)
          #OR
dpois(x=0, lambda = 15)

```

#### 1c. Pr(X=3) in one session

```{r}
dpois(x=3, lambda = 15)

```
#### 1d. Plot Poisson mass function with 0 &le; x &le; 30

```{r}
titi <- dpois(x=0:30, lambda = 15)
barplot(titi, names.arg = 0:30, space = 0, xlab = "Number of calls heard", ylab = "Probability of calls heard", main = "Probability Mass Function")

```

#### 1e. Simulate 104 results from this distribution.

```{r}
two_yrs <- rpois(n=104, lambda = 15)

```

#### 1f. Plot simulation using hist() and xlim(), 0-30. Compare to 1d.

```{r}
par(1, 2)


hist(two_yrs, probability = TRUE, xlim = c(0,30), main = "Two Years of Field Work", xlab = "Number of titi calls heard", ylab = "Probability of hearing calls")

titi <- dpois(x=0:30, lambda = 15)
barplot(titi, names.arg = 0:30, space = 0, xlab = "Number of calls heard", ylab = "Probability of calls heard", main = "Probability Mass Function")

```

### CHALLENGE 2:
*Zombie apocalypse survivors*
```{r}
library(curl)
z <- curl("https://raw.githubusercontent.com/difiore/ADA-2019/master/zombies.csv")

dZ <- read.csv(z, header = TRUE, sep = ",", stringsAsFactors = FALSE)

head(dZ)

```


#### 2a. **Population** mean and standard deviation for each quantitative random variable (height, weight, age, number of zombies killed, years of education).qrt

```{r}

(muH <- mean(dZ$height))
(siH <- sqrt(sum((dZ$height - muH)^2)/1000))
(muW <- mean(dZ$weight))
(siW <- sqrt(sum((dZ$weight-muW)^2)/1000))
(muA <- mean(dZ$age))
(siA <- sqrt(sum((dZ$age-muA)^2)/1000))
(muZK <- mean(dZ$zombies_killed))
(siZK <- sqrt(sum((dZ$zombies_killed-muZK)^2)/1000))
(muEd <- mean(dZ$years_of_education))
(siEd <- sqrt(sum((dZ$years_of_education-muEd)^2)/1000))

```
#### 2b. Use {ggplot} to make boxplots of each variable by gender.

```{r}
library(ggplot2)
library(gridExtra)

H <- ggplot(dZ, aes(x= gender, y= height)) + geom_boxplot() + ggtitle("Survivor Height") + xlab("Gender") + ylab("Height")


W <- ggplot(dZ, aes(x= gender, y= weight)) + geom_boxplot() + ggtitle("Survivor Weight") + xlab("Gender") + ylab("Weight")


A <- ggplot(dZ, aes(x= gender, y= age)) + geom_boxplot() + ggtitle("Survivor Age") + xlab("Gender") + ylab("Age")


ZK <- ggplot(dZ, aes(x= gender, y= zombies_killed)) + geom_boxplot() + ggtitle("Survivor's Kill Count") + xlab("Gender") + ylab("Zombies Killed")


Ed <- ggplot(dZ, aes(x= gender, y= years_of_education)) + geom_boxplot() + ggtitle("Survivor Education") + xlab("Gender") + ylab("Years of Education")

grid.arrange(H, W, A, ZK, Ed, ncol=3, nrow=2)

```

#### 2c. Use {ggplot} to make scatterplots of height and weight in relation to age. Are they related? In what way?

```{r}
H.A <- ggplot(dZ, aes(x=age, y=height)) + geom_point() + ggtitle("Survivor Height by Age") + xlab("Age") + ylab("Height") 

W.A <- ggplot(dZ, aes(x=age, y=weight)) +geom_point() + ggtitle("Survivor Weight by Age") + xlab("Age") + ylab("Weight") 

grid.arrange(H.A, W.A, ncol=2)

```
  * Both scatterplots appear to have a positive linear correlation, therefore both height and weight are correlated to age. This makes sense since both height and weight increase with development over time. However, weight is not as strongly correlated since resource acquisition can depend on multiple variables, such as skill level or resource availability. 

#### 2d. Determine which quantitative variables are drawn from normal distribution using histograms and Q-Q plots. 
  * for those that are not, determine what distribution they're drawn from.
  
```{r}
attach(dZ)

par(mfrow= c(1, 2))

qqH <- {
  qqnorm(height, main = "Normal QQ Plot- Height")
  qqline(height, col="red") } 

histH <- hist(height, main = "Histogram of Survivor Height", probability = TRUE)

qqW <- {
  qqnorm(weight, main = "Normal QQ Plot- Weight")
  qqline(weight, col="red") }

histW <- hist(weight, main = "Histogram of Survivor Weight", probability = TRUE)

qqA <- {
  qqnorm(age, main = "Normal QQ Plot- Age")
  qqline(age, col="red") }

histA <-  hist(age, main = "Histogram of Survivor Age", probability = TRUE)

qqZK <- {
  qqnorm(zombies_killed, main = "Normal QQ Plot- \nZombies Killed")
  qqline(zombies_killed, col="red") }

histZK <- hist(zombies_killed, main = "Histogram of Number \nof Zombie Kills", probability = TRUE)

qqEd <- {
  qqnorm(years_of_education, main = "Normal QQ Plot- \nYears of Education")
  qqline(years_of_education, col="red") }

histEd <- hist(years_of_education, main = "Histogram of Survivors' \nYears of Education", probability = TRUE)

```
    * For the continuous variables of height, weight, and age, most of the observations do follow a normal distribution, with only the tails deviating from normal. 
    * Since the zombies killed and age variables are discrete variables, they cannot be normally distributed, as clearly seen in the QQ plots and histograms; both factors are heavily skewed to the left. 
      - Assuming these survivors use the double-tap rule, zombies are either dead dead or living dead, with a continuous amount of time to be either of these options, so a poisson distribution will fit these data well.
      - While partial years of education are possible, this data set appear to only consit of whole years. With only a set number of "trials," education is best fit by a binomial distribution.
    
#### 2e. Create a single subset of 30 survivors and calculate mean/sd for each variable. Estimate standard error for each variable and construct 95% confidence intervals for each mean. 

```{r}
library(matrixStats)
set.seed(1)

alexandria <- dZ[sample(nrow(dZ), size=30, replace=FALSE),]

se <- function(x) {
  sd(x)/sqrt(length(x))
}

ci.H <- function(column, data=NULL) {
  column <- eval(substitute(column), data, parent.frame())
  qnorm(0.95/2, mean = mean(column), sd = sd(column))
}

ci.L <- function(column, data=NULL) {
  column <- eval(substitute(column), data, parent.frame())
  qnorm(0.05/2, mean = mean(column), sd = sd(column))
}

a.m <- colMeans(alexandria[,c(5:8,10)])
a.sd <- apply(alexandria[,c(5:8,10)],2,sd)
a.se <- apply(alexandria[,c(5:8,10)],2,se)
a.ci.H <- apply(alexandria[,c(5:8,10)],2,ci.H)
a.ci.L <- apply(alexandria[,c(5:8,10)],2,ci.L)

alex_Stats <- matrix(c(a.m, a.sd, a.se, a.ci.H, a.ci.L),5,5,byrow = FALSE)
colnames(alex_Stats) <- c("Mean", "Sd", "Sd.Er.", "CI-High", "CI-Low")
rownames(alex_Stats) <- c("Height", "Weight", "Age", "Zom.Killed", "Yrs Ed")
alex_Stats

```

#### 2f. Repeat the previous step to get 100 total samples. Calculate mean/sd of sampling distributions for each variable and compare to mean/sd in *2e*. Describe distributions.

