---
title: "davis-allison-ADA-homework-2"
author: "Allison Davis"
date: "February 20, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## MODULE 11
#### Probabilities and Distributions

### CHALLENGE 1:
*X = no. calls heard in each session;* *expected mean = lambda = 15*

#### 1a. Pr(X>8) in one session

``` {r}
1- ppois(q=8, lambda = 15)

```

#### 1b. Pr(X=0) in one session

```{r}
ppois(q=0, lambda = 15)
          #OR
dpois(x=0, lambda = 15)

```

#### 1c. Pr(X=3) in one session

```{r}
dpois(x=3, lambda = 15)

```
#### 1d. Plot Poisson mass function with 0 &le; x &le; 30

```{r}
titi <- dpois(x=0:30, lambda = 15)
barplot(titi, names.arg = 0:30, space = 0, xlab = "Number of calls heard", ylab = "Probability of calls heard", main = "Probability Mass Function")

```

#### 1e. Simulate 104 results from this distribution.

```{r}
two_yrs <- rpois(n=104, lambda = 15)

```

#### 1f. Plot simulation using hist() and xlim(), 0-30. Compare to 1d.

```{r}
par(1, 2)


hist(two_yrs, probability = TRUE, main = "Two Years of Field Work", xlab = "Number of titi calls heard", ylab = "Probability of hearing calls") + xlim(0, 30)

titi <- dpois(x=0:30, lambda = 15)
barplot(titi, names.arg = 0:30, space = 0, xlab = "Number of calls heard", ylab = "Probability of calls heard", main = "Probability Mass Function")

```

### CHALLENGE 2:
*Zombie apocalypse survivors*

#### 2a. **Population** mean and standard deviation for each quantitative random variable (height, weight, age, number of zombies killed, years of education).

#### 2b. Use {ggplot} to make boxplots of each variable by gender.

#### 2c. Use {ggplot} to make scatterplots of height and weight in relation to age. Are they related? In what way?

#### 2d. Determine which quantitative variables are drawn from normal distribution using histograms and Q-Q plots. 
  * for those that are not, determine what distribution they're drawn from.
  
#### 2e. 

